---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# SamsBeautifulPackage

The goal of SamsBeautifulPackage is to offer R users several ready-made functions for data cleaning and wrangling. Previously a user had to use several lines of code which our package can now accomplish with one simple function. We hope that you find this useful!

## Installation

To install the development version from [GitHub](https://github.com/) use:

``` r
# install.packages("remotes")
remotes::install_github("jonasdfreeman/SamsBeautifulPackage")
```

## General Usage

For examples of how to use the functions in this package, the user should direct their attention to the `Functions In SamsBeautifulPackage` vignette. This vignette offers the user a series of examples of how to apply the functions in the package to a dummy dataset which contains several common issues in real-world data.

## User Guide for Mouse Data

This package was primarily designed to deal with the `mousedata.xlsx` file which was presented in Statistics 108: Introduction to R (taught by James Xenakis). We have included a helpful vignette entitled `Using SamsBeautifulPackage On Mouse Data` which goes through a step by step process outlining how our functions can be used to analyze mouse data.

## Key Highlights:

1. **`check_duplicates(df)`**: Checks for and displays duplicate rows.
2. **`check_values(data, column_name, value1, value2)`**: Verifies if a column has only specific valid values.
3. **`find_non_numeric(data, column_name)`**: Coerces a column to numeric data. Identifies and removes rows with non-numeric values in a column.
4. **`generate_plot(data, x_var, y_var)`**: Generates scatter plots of two variables.
5. **`percent_diff(df, col1, col2)`**: Identifies greater than 20% differences between two columns.
6. **`remove_zeroes(data, column_name)`**: Removes rows where a column has zero, empty, or NA values.
7. **`data_summary(data)`**: Gives the user a snapshot with some summary statistics of the numeric data.

### Contents

SamsBeautifulPackage contains one dataset called `ourdata`:

```{r showdata}
library(SamsBeautifulPackage)
data(package = 'SamsBeautifulPackage')
```

This dataset is originally sourced from Kaggle under the CC0 1.0 Universal licence. This is data that confirmed the grade of performance with age and some exercise performance data.

Here is a link to the original dataset source: https://www.kaggle.com/datasets/kukuroo3/body-performance-data

```{r}
head(ourdata)
``` 

The dataset was comrpised of 13393 rows and 12 columns, but the data in the package has been wrangled to just 13392 rows (see data-raw folder).

```{r}
str(ourdata)
```

### Function Examples

For more information on the specific functionality of the functions in SamsBeautifulPackage, please see our documentation by typing `help(package = "SamsBeautifulPackage")`.

Here are some fun examples of how our package can be used:

```{r}
# Generate a plot automatically by inputting arguments into the generate_plot() function
generate_plot(ourdata, "age", "gripForce")
```


```{r}
# Check whether everyone's gender is either "M" or "F"
check_values(ourdata, "gender", "M", "F")
```


```{r}
# Here is a cool plot which plots heigh vs weight in one simple function
generate_plot(ourdata, "height_cm", "weight_kg")
```



## Citation

To cite SamsBeautifulPackage package, please use:

```{r}
citation("SamsBeautifulPackage")
```

## License

Data are available by  [CC-0](https://creativecommons.org/share-your-work/public-domain/cc0/)

## Additional data use information

Anyone interested in publishing the data should contact Sam Lindemann samlindemann@college.harvard.edu

